---
title: "COMPSCIX 415.2 Homework 4"
author: "Aparna Vaidya"
date: "2/22/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


My Github repository for my assignments can be found at this URL <https://github.com/aparnav99/compscix-415-2-assignments.git>


```{r load_packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(nycflights13)
```

###5.6.7 Exercises

__2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).__

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
```

```{r}
not_cancelled %>% count(dest)

not_cancelled %>% group_by(dest) %>% summarize(n = n())
```

```{r}
not_cancelled %>% count(tailnum, wt = distance)

not_cancelled %>% group_by(tailnum) %>% summarize(wt = sum(distance))
```


__4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?__

There seems to be a postive relationship between average delay and number of flights cancelled. Greater the delay, higher the number of flights cancelled.

```{r}
flights %>%
  group_by(day) %>%
  summarise(cancelled = mean(is.na(dep_delay)),
            mean_dep = mean(dep_delay, na.rm = T),
            mean_arr = mean(arr_delay, na.rm = T)) %>%
  ggplot(aes(y = cancelled)) +
  geom_point(aes(x = mean_dep), colour = "red") +
  geom_point(aes(x = mean_arr), colour = "blue") +
  labs(x = "Avg delay per day", y = "Number of cancelled flights per day")
```


__6. What does the sort argument to count() do. When might you use it?__
sort argument is used to sort the column by count.

```{r}
flights %>%
  count(carrier, sort = TRUE)
```


### 10.5 Exercises

__1.How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).__

Tibble shows column types along with column names. Also it displays first 10 rows.

```{r}
mtcars

mtcars %>% as.tibble()
```


__2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?__

Data frame returns results even when there is a partial match on column names which can be misleading. Tibble will try an exact match on column names.

Data frames returns a vector if only one column is passed as input and it retuns a data frame if multiple columns are passed. This makes it difficult to handle results. Tibble always retuns a tibble as resulting type.

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]


tb <- df %>% as.tibble()

tb$x
tb[, "xyz"]
tb[, c("abc", "xyz")]

```


__3.If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?__

```{r}
var <- "xyz"
df[[var]]


```


__6. What option controls how many additional column names are printed at the footer of a tibble?__

tibble.max_extra_cols


###12.3.3 Exercises

__2. Why does this code fail?__
__table4a %>% __
__  gather(1999, 2000, key = "year", value = "cases")__
__#> Error in inds_combine(.vars, ind_list): Position must be between 0 and n__

Because 1999 and 2000 are non-syntactic names(they don't start with a letter). They need to be enclosed in backticks.

```{r}
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```


__3. Why does spreading this tibble fail? How could you add a new column to fix the problem?__

__people <- tribble(__
__  ~name,             ~key,    ~value,__
__  #-----------------|--------|------__
__  "Phillip Woods",   "age",       45,__
__  "Phillip Woods",   "height",   186,__
__  "Phillip Woods",   "age",       50,__
__ "Jessica Cordero", "age",       37,__
__  "Jessica Cordero", "height",   156__
__)__

Spreading this tibble will fail as the 'age' key is not unique per observation. Using distinct for rows will help.

```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

people %>% distinct(name, key, .keep_all = TRUE) %>% spread(key, value)

```


__4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?__

__preg <- tribble(__
__  ~pregnant, ~male, ~female,__
__  "yes",     NA,    10,__
__  "no",      20,    12__
__)__

```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

preg %>% 
  gather(`male`, `female`, key = "gender", value = "count", na.rm = TRUE)
```


###12.4.3 Exercises

__1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.__

extra argument specifies what to do if there are extra values.
fill argument specifies what to do if there are missing values.

```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "merge")

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "right")
```


__2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?__

Remove argument removes the input columns from the resulting dataframe and creates a new column. Setting it to FALSE will retain the old columns and create new columns in resulting datarame.


### Import and Export Data

```{r}
baby_names <- read.delim("/Users/aparnavaidya/Downloads/UCB-Intro To Data Science/compscix-415-2-assignments/Lab4/baby_names.txt", sep = "|")

glimpse(baby_names)

write_csv(baby_names, "/Users/aparnavaidya/Downloads/UCB-Intro To Data Science/compscix-415-2-assignments/Lab4/baby_names.csv")

baby_names_csv <- read.csv("/Users/aparnavaidya/Downloads/UCB-Intro To Data Science/compscix-415-2-assignments/Lab4/baby_names.csv")

glimpse(baby_names_csv)

```

